(tentative) Rotational Spectral Unmixing of Exoplanets

### 2016-08-08 paper/main.tex
A figure of 4 EPOXI observations is added. A figure of retrieved result from EPOXI June observation is added for a test. Text updated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Algebraic Model}
\label{ss:model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Assuming that the planetary surface is Lambertian scatterer everywhere, with a certain number $K$ of mutually distinct surface types, 
the disk-integrated spectra can be written as a weighted summation of the reflectance spectra of different surface types, e.g., 
%%%
\begin{equation}
d_{ij} = R_p^2 \sum _{i,k} f_{ik}^{\ast } \, s_{kj}
\end{equation}
%%%
where $R_p$ is the planetary radius, $i$, $j$, and $k$ are indices for the observation epochs, bands, and the surface types, respectively. 
The symbols represent as follows:
$d_{ij}$ is the observed apparent albedo (``$d$'' for data) at $i$-th epoch and $j$-th band, 
$f_{lk}^{\ast }$ is the contribution factor of the $k$-th surface type at $i$-th observational epoch, and 
$s_{kj}$ is the reflectance spectra of $k$-th surface type at $j$-th band.

The contribution factor $f^{\ast }$ is the weighted summation of the area fraction of $k$-th surface type, e.g., 
%%%
\begin{equation}
f^{\ast }_{ik} = \sum _{l} W_{il} f_{lk}
\end{equation}
%%%
where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type at $l$-th longitudinal slice, and 
$W_{il}$ is the weight function for $i$-th epoch and $l$-th longitudinal slice which depends only on the observational geometry. 
As a result,
%%%
\begin{equation}
d_{ij} = R_p^2 \sum _{l,k} W_{il} \, f_{lk} \, s_{kj}
\end{equation}
%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Baysean Formulation and Regularization}
\label{ss:regularization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to 4 EPOXI observation of the Earth}
\label{ss:epoxi}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Algebraic Formulation}
\label{ss:model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Assuming that the planetary surface is Lambertian scatterer everywhere, with a certain number $K$ of mutually distinct surface types, 
the disk-integrated spectra can be written as a weighted summation of the reflectance spectra of different surface types, e.g., 
%%%
\begin{equation}
d_{ij} = R_p^2 \sum _{i,k} f_{ik}^{\ast } \, s_{kj}
\end{equation}
%%%
where $R_p$ is the planetary radius, $i$, $j$, and $k$ are indices for the observation epochs, bands, and the surface types, respectively. 
The symbols represent as follows:
$d_{ij}$ is the observed apparent albedo (``$d$'' for data) at $i$-th epoch and $j$-th band, 
$f_{lk}^{\ast }$ is the contribution factor of the $k$-th surface type at $i$-th observational epoch, and 
$s_{kj}$ is the reflectance spectra of $k$-th surface type at $j$-th band. 
The maximum number of $i$, $j$ and $k$ will be denoted by $I$, $J$, and $K$, below.

The contribution factor $f^{\ast }$ is the weighted summation of the area fraction of $k$-th surface type, e.g., 
%%%
\begin{equation}
f^{\ast }_{ik} = \sum _{l} W_{il} f_{lk}
\end{equation}
%%%
where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type at $l$-th longitudinal slice, and 
$W_{il}$ is the weight function for $i$-th epoch and $l$-th longitudinal slice which depends only on the observational geometry. 
The maximum number of $l$ is denoted by $L$. 
As a result,
%%%
\begin{equation}
d_{ij} = R_p^2 \sum _{l,k} W_{il} \, f_{lk} \, s_{kj}
\end{equation}
%%%

@@

%%%
\begin{equation}
\tilde d_{ij} = \sum _{i,k} f_{ik}^{\ast } \, s_{kj}
\end{equation}
%%%
where $\tilde d_{ij} = d_{ij}/R_p^2$. 
Without any constraints on $f^{\ast }$ and $s$, it is trivial to notice the degeneracy in the solutions. 
Using an arbitrary regular matrix $X$ with rank $K$, a particular set of solution $\{ f^{\ast },\,s\}$ implies another set of solution $\{ f^{\ast }X^{-1},\,Xs\}$.

Now that they are subject to the constraints (equation (\ref{eq:constraints_ast})), the range of acceptable solution is limited. 
However, the constraints cannot be sufficient to reduce them to a unique solution. 
One can create a infinite number of candidates for $f_{ik}^{\ast }$ which satisfies both of the two constraints for $f$, then $s$ may be obtained by $s = f^{+} \tilde d$, where $f^+$ is the pseudo-inverse matrix (note that $K < I$ to be able to find the solution) and among these solutions there should be many that satisfy the constraint for $s$ as well. 
\memoYF{Need to brush up. Possibly need to consider SVD of $f$. }

To see this, we demonstrate the application to the mock data with known solution in Section \ref{s:mockdata}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bayesian Formulation}
\label{ss:regularization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%
\begin{eqnarray}
Posterior = \prod\limits_{i,j} \frac{1}{\sqrt{2\pi }\sigma_{ij}} &&  \exp \left[ - \frac{(d_{ij} - \sum _{lk} W_{il} f_{lk} s_{kj})^2}{2 \sigma _{ij}^2} \right] \notag \\
&& \times \; \mathcal{P} (\{f_{l,k}\}) \cdot \mathcal{P}(\{s_{l,k}\}) 
\end{eqnarray}
%%%
where $\mathcal{P} (X) $ is a prior for $X$.

In order to take account of constraints (equation (\ref{eq:constraints})), we adopt the following transformation of variables: 
%%%
\begin{eqnarray}
\begin{cases}
t_{kj} = \displaystyle \ln \left[ \frac{ s_{kj} }{1 - s_{kj} } \right]  \;\;\; & \mbox{for any $k$, $j$} \\
g_{lk} = \displaystyle \ln \left[ \frac{ f_{lk} }{1 - \sum _{m=0}^{k} f_{lm} } \right] \;\;\; & \mbox{for $k=1, ..., K-1$, any $j$} 
\end{cases}
\end{eqnarray}
%%%
Note that there is a 1-to-1 correspondence between $t$ and $s$ or $g$ and $f$ where $s$ and $f$ satisfy our constraints.

@@

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to 4 EPOXI observations of the Earth}
\label{ss:epoxi}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!bt]
    \begin{center}
\includegraphics[width=\hsize]{EPOXI_vislightcurve_4obs.pdf}
    \end{center}
    \caption{4 light curves of the Earth obtained with EPOXI mission.}
\label{fig:EPOXIlc}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!bt]
    \begin{center}
    \includegraphics[width=\hsize]{June_xmed_std_GPReg.pdf}
    \end{center}
    \caption{Result of inversion of EPOXI June data. \memoYF{to be revised}}
\label{fig:mcmc_June_tmp}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
